### 2023-07-26 thesis/sections/3-exp-setup.tex
captioning game
description language games

The language games in this research have an asymmetric setup.
One agent, the sender is shown some information and needs to generate a message.
This message is received by the second agent, the receiver.
The receiver needs to parse this message and combine it with the information they are given as well.
The receiver then makes a prediction, which is compared to the ground truth.
The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
All weights of the agents are adapted in the same way.
\cmtDK{Gumbel-Softmax vs REINFORCE}
\cmtDK{better in theoretical background?}
\cmtDK{vocabulary?}

All language games are set up and run in the EGG framework \citep{Kharitonov2019}, developed in PyTorch.
This framework consists of heavy configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
Furthermore, runs of games can be saved to analyze the used messages of the agents and how they change over the duration of the learning.

@@

@@

@@

\paragraph{Caption generators:}

\paragraph{Coordinate predictors:}

>>>>

The language games in this research have an asymmetric setup.
One agent, the sender is shown some information and needs to generate a message.
This message is received by the second agent, the receiver.
The receiver needs to parse this message and combine it with the information they are given as well.
The receiver then makes a prediction, which is compared to the ground truth.
The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
All weights of the agents are adapted in the same way.
\cmtDK{Gumbel-Softmax vs REINFORCE}
\cmtDK{better in theoretical background?}
\cmtDK{vocabulary?}
\cmtDK{batch iterator}
All language games are set up and run in the EGG framework \citep{Kharitonov2019}, developed in PyTorch.
This framework consists of heavy configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
Furthermore, runs of games can be saved to analyze the used messages of the agents and how they change over the duration of the learning.

@@

\cmtDK{always ResNet-3}

@@

\cmtDK{look at emerged language: distribution of vocab/ngrams, how do they relate to the images (clustering?)}

@@

\paragraph{Caption generators:}
In a next step, it is tested if the agents can learn to extract features of the objects together.
For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
The receiver only sees the image without additional information.
\cmtDK{human bias vs emergent language}

This setup is based on the single neural model described above.
The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.

The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
The image encoders are set up as described in \citet{Johnson2017}.
A feature extractor extracts basic features of the images.
Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
This size was found to have the best results during the experiments.
The two resulting encoded images are flattened and concatenated.
A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.

The receiver uses the same architecture to encode the original image as the sender.
The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
The loss is calculated with cross entropy.

Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.

During the experiments, the following variables are adjusted, and the results are compared:
(1) the vocabulary size $|V|$,
(2) the LSTM hidden size of the sender wrapper $h_s$ and
(3) the LSTM hidden size of the receiver wrapper $h_r$.
The same metrics as in the pre-experiments are used to evaluate the results.

\paragraph{Coordinate predictors:}
In the final experiments, it is tried to eliminate as much human knowledge as possible.
For that reason, the agents are tasked to only predict the location of the target object.
This task is approached in two steps.
In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
The receiver doesn't come in contact with any human knowledge, not as input nor as output.
This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
In the second step, the caption is also removed from the sender.
Instead, a masked image points the sender towards the target object.
In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.

The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
The sender encodes the original image the same as in the previous paragraph.
In the first setup, the description of the target object is encoded, using an LSTM.
For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
These embeddings are the input for the LSTM.
The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
Choosing a size of 10 for both variables proved to give good results in the experiments.
In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
The resulting vector is passed to the sender wrapper LSTM, to generate the message.

For the second approach, the masked image is passed only through a feature extractor.
Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.

The receiver contains of two parts.
First, the original image is encoded with an image encode of the same setup.
This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
The resulting vector is the passed through the second part of the receiver, the predictor.
This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
A \emph{ReLU} function is applied in between.
This setup is trained to combine the important information from both the image and the message of the sender.
The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.

As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.

During the experiments, the effects of the following variables are compared:
(1) the vocabulary size $|V|$,
(2) the LSTM hidden size of the sender wrapper $h_s$,
(3) the LSTM hidden size of the receiver wrapper $h_r$ and
(4) the embedding size of the sender $e_s$.
